---
title: "Plotting a categorical x and numeric y"
output:
  xaringan::moon_reader:
    highlightLines: false
    lib_dir: libs
    css: xaringan-themer.css
    nature: 
      countIncrementalSlides: false
      highlightStyle: github
      highlightLines: true
      ratio: 16:10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 8,fig.height = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 2)
library(flipbookr)
library(xaringan)
library(xaringanthemer)
library(tidyverse)

ril_link <- "https://raw.githubusercontent.com/ybrandvain/datasets/refs/heads/master/clarkia_rils.csv"
ril_data <- readr::read_csv(ril_link) |>
  dplyr::mutate(growth_rate = case_when(growth_rate =="1.8O" ~ "1.80",
                                          .default = growth_rate),  
                growth_rate = as.numeric(growth_rate),
                visited = mean_visits > 0)
gc_rils <- ril_data |>
  filter(location == "GC", !is.na(prop_hybrid), ! is.na(mean_visits))|>
  select(petal_color, petal_area_mm, num_hybrid, offspring_genotyped, prop_hybrid, mean_visits , asd_mm,visited )
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
xaringanthemer::style_duo_accent(
  primary_color = "#ed62ac",
  secondary_color = "#3A5F0B",text_color = "#3A5F0B",title_slide_text_color = "#3A5F0B",
  header_font_google = google_font("Lato"),
  code_font_google   = google_font("Roboto Mono"),
  base_font_size = "35px",
  code_font_size = ".6rem",
  text_slide_number_font_size = ".6em")

# Set a global ggplot2 theme
theme_set(
 theme_xaringan( text_font_size = 35  )
)

```

```{r xaringanExtra-share-again, echo=FALSE}
xaringanExtra::use_share_again()
# adds share bar to flipbooks
```



```{css, echo=FALSE}
.left-column, .right-column {
  display: inline-block;
  vertical-align: top;
  width: 46%;
  padding: 0 1%;
}

.left-column pre {
  font-size: 85%;
  max-height: 400px;
  overflow-y: auto;
}
```

**Plotting a numerical y and a categorical x is hard**  

This slideshow  works you through challenges and how to overcome them! These solutions include:  

- Using jitter to avoid overplotting.  
- Using color to differentiate values.  
- Suppressing a redundant legend.  
- Showing means on a plot.  

---

**Step 1: Basic ggplot setup**   
Setting up aesthetic mapping <br> <br>

.left-column[
```{r, eval=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid))
```
]

.right-column[
```{r,echo=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid))
```
]


---

**Step 1: Basic ggplot setup**  
Adding points <br> <br>


.left-column[
```{r, eval=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid)) +
  geom_point()
```
]

.right-column[
```{r,echo=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid))+
  geom_point()
```
]



---

**Step 1: Basic ggplot setup**  
OH NO those points are too small.  
Let's try bigger, somewhat transparent points.


.left-column[
```{r, eval=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid)) +
  geom_point(size = 6, 
             alpha = .3)
```
]

.right-column[
```{r,echo=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid)) +
  geom_point(size = 6, 
             alpha = .3)
```
]


---



class: center middle section-slide

# Try jitterplots  



---

**Step 2: Try jitterplots**  
It's still tough to differentiate overlapping points. 
Let's switch to jitterplots to spread out our x.


.left-column[
```{r, eval=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)
# I set height equal to zero
# So the plot doesn't lie.
```
]

.right-column[
```{r,echo=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)
# I set height equal to zero
# So the plot doesn't lie.
```
]


---


**Step 2: Try jitterplots**  
That helps - but now categories blend :(  
Lets redundantly map visited on to color.


.left-column[
```{r, eval=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid,
             color = visited)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)
```
]

.right-column[
```{r,echo=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid,
             color = visited)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)
```
]


---


**Step 2: Try jitterplots**  
Now the unnecessary legend takes up too much space.  
Let's remove it!


.left-column[
```{r, eval=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid,
             color = visited)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)+
  theme(legend.position = "none")
```
]

.right-column[
```{r,echo=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid,
             color = visited)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)+
  theme(legend.position = "none")
```
]


---




class: center middle section-slide

# Add summaries 



---


**Step 3: Add summaries**  
Now let's add each group's mean <br><br>



.left-column[
```{r, eval=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid,
             color = visited)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)+
  theme(legend.position = "none")+
  stat_summary(fun = "mean")
```
]

.right-column[
```{r,echo=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid,
             color = visited)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)+
  theme(legend.position = "none")+
  stat_summary(fun = "mean")
```
]

---


**Step 3: Add summaries**  
Those points are tiny.   
Let's make them big and black



.left-column[
```{r, eval=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid,
             color = visited)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)+
  theme(legend.position = "none")+
  stat_summary(fun = "mean", 
               color = "black", 
               size = 4)
```
]

.right-column[
```{r,echo=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid,
             color = visited)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)+
  theme(legend.position = "none")+
  stat_summary(fun = "mean", 
               color = "black", 
               size = 1.5)
```
]

---



**Step 3: Add summaries**  
Those points are tiny.   
Let's make them big and black



.left-column[
```{r, eval=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid,
             color = visited)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)+
  theme(legend.position = "none")+
  stat_summary(fun = "mean", 
               color = "black", 
               size = 1.5)+
  stat_summary(aes(group = 1),
    color = "black", geom = "line", 
    lty = 2, size = 2)
```
]

.right-column[
```{r,echo=FALSE}
gc_rils|>
  ggplot(aes(x = visited,
             y = prop_hybrid,
             color = visited)) +
  geom_jitter(size = 6,  
             alpha = .3, 
             height = 0)+
  theme(legend.position = "none")+
  stat_summary(fun = "mean", 
               color = "black", 
               size = 1.5)+
  stat_summary(aes(group = 1),
    color = "black", geom = "line", 
    lty = 2, size = 2)
```
]

---



class: center middle section-slide

# *'Fin'*


